\subsection{Research questions}\label{sec:rqs}
Beyond the overall goal of analyzing the state of the art in traceability research in software/systems engineering and identifying its open research challenges, we are especially interested in the following subtopics that will be the main focus of our analysis when evaluating the papers resulting from the survey.

\subsubsection{RQ1: How are traces represented?}
\label{sec:rq1}
%\nb{How does traceability work}{Common features in the traceability landscape}
%We investigated work on traceability through a survey of papers that mention explicitly traceability and modelling. Not all features are covered the same and we further precise our feature model with the results of this survey to enlighten current limitations - and thus opportunities for future work. This study forms the basic knowledge to build a feature model describing what we consider the main characteristics of traceability methodologies. 
We are interested in knowing what languages are used to represent trace and how expressive they are. E.g. is it possible to define types of traces? And use them to link clusters of artefacts (or just individual elements)? Can we express the confidence we have in those traces?


\subsubsection{RQ2: How are traces identified?}
\label{sec:rq2}
%\nb{How does trace identification work}{Technical details of identification features}
Do traces need to be manually created or can be derived from existing information? If so, what methods can be used for such derivation?. These are some of the topic we would like to answer in this RQ.
%AI for identification, domain contextualization, topic modelling, there is a bunch of work. What trends tend to follow work on AI-assisted trace link identification ?

\subsubsection{RQ3: What tool support is available to manage and apply traces?}
\label{sec:rq3}
%\nb{Why is traceability important}{Broader areas benefiting from traceability}
Finally we want to draw a general map of the techniques that traceability approaches use to store and maintain trace artefacts of quality.
%Where is traceability used right now? Among what types of artefacts and in order to help in what tasks?


