\section{Tracea -- a DSL for traceability}
A DSL\footnote{We use the DSL term to refer to Domain-Specific Modeling Languages, which is the subset of DSLs that are relevant to us} is defined through three main components \cite{kleppe2008-DSLs-with-metamodels}: abstract syntax, concrete syntax, and semantics. 

The abstract syntax defines both the language concepts and their relationships, and also includes well-formedness rules constraining the models that can be created. Metamodeling techniques are normally used to define the abstract syntax. The concrete syntax defines a notation (textual, graphical or hybrid) for the concepts in the abstract syntax, and a translational approach is normally used to provide semantics, though most of the time it is not explicitly formalized.

The metamodel definition directly includes some of the terminology discussed in the previous section. For other concepts we have made sure the DSL is complete enough to allow for that kind of processing or analysis. 
For instance, the concept of traces transcribes directly into classes of the metamodel (\textit{e;g.,} traces, links, artefacts). Processes span over subsets of classes (\textit{e.g.,} granularity and typing).
Characteristics such as vertical traceability, or implicit traces do not appear directly in the metamodel but are either found disseminated among its structural composition or the DSL elements allow for their implementation in specific use cases. 

In this section,
\begin{itemize}
    \item We provide a running example: transclusion for traceability between certification documents and design models.
    \item We present the metamodel, decomposed into five sub-areas of traceability and detail each of them. We reuse the example to show the soundness of our choices.
    \item We provide a textual syntax in the form of a {JSon-based grammar}. 
\end{itemize}

\input{sections/mm-0example}

\subsection{Metamodel} 

\input{sections/mm-1core}
\input{sections/mm-2granularity}
\input{sections/mm-3relationship}
\input{sections/mm-4integrity}
\input{sections/mm-5textsyntax}


\subsection{Conclusion and future work}
In this deliverable, we have developed a DSML for traceability. 
In the future, this DSL could be extended with the consideration of belief uncertainty as addressed by Loli Burgueno in the Modelia project~\cite{burgueno2019-uncertainty}. Traceability has an important human factor to consider. This extension would allow reasoning, manipulation and decision making with belief and subjective opinions of the referees involved in the process.

Then, since time plays a fundamental role in the decline of coherence in a system, this dimension shall be investigated meticulously as well. Coordinating the record of tracing operation time wise would give the opportunity to evaluate the aging of traces. In this first version, we simply record \textit{timestamps} for all trusteable artefacts. These stamps shall be augmented with \textit{TemporalEMF}, a profile for temporal metamodelling developed in SOM specifically for EMF solutions~\cite{gomez2018-temporalEMF}.  

Finally, another interesting follow up would be to adapt this DSL to UML, SysML or DOORS in order to populate it with real world traces. This future work requires the evaluation of entry points from Tracea to other languages. It also requires a definition of the purpose of tracing in the projects targeted.

